Field and experimental data together with computational models reveal how cleaner fish adjust decisions in a biological market

### 2022-03-08 manuscript_1.0/manuscript_1.0.tex
Updated manuscript tp the PLOS Comp. Biol. Updated figures with new figures.

Over the last decade, our team tested over a hundred wild-caught
cleaner fish in the exact same paradigm of the ephemeral reward
task \citep{salwiczek_Adult_2012, wismer_Variation_2014, triki_Decrease_2018, triki_Biological_2019, triki_Brain_2020}. These fish often come from
different reef locations around the study field site at Lizard Island,
Great Barrier Reef in Australia. Further investigation of the local
eco-sociological conditions revealed that cleaner and client
fish population densities have a substantial impact on cleaner fish
performance in the task. Cleaner fish from reef sites with relatively
low densities were more likely to fail at solving the task \citep{triki_Biological_2019, triki_Decrease_2018, wismer_Variation_2014}. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply
low supply and high demand for cleaning services.
Under such conditions, there are fewer occasions under
which failing to give priority to a visitor translates to an empty
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities
are less likely to exert their partner choice options
and are hence more likely to wait for inspection if not given cleaning
priority \citep{triki_Brain_2020}. In the absence of partner choice
cleaner fish should not give priority to visitors anymore.

@@

The empirical data were collected between 2010 and 2019 always during
the austral winter months June to August from a total of five
study reef sites (Corner Beach-CB, Horseshoe-HS, Mermaid Cove-MC,
Northern Horseshoe-NHS, and The Crest-TC) at Lizard Island
(\(14.6682° S, 145.4604° E\)), Great Barrier Reef, Australia.
The data consist of three sets: fish censuses, field
observations of cleaner-client interactions to quantify the probability
of visitors leaving if made to wait, and the
performance of wild-caught cleaner fish in the ephemeral reward test.
In total, we have twelve site/year data sets for fish censuses and
corresponding performance in the lab test. Thus, some sites were sampled more
than once. To estimate the population density of cleaner fish
and their clients at a given site in
a given year, Triki \emph{et al.} \citep{triki_Biological_2019} used a series of ten
transects of \(30m\) each. Observers swam along the transect lines placed on
the reef and first counted the visible large-bodied adult fish
(species with total length \(TL \geq 10cm\)) including cleaner fish
on a width of \(5m\), and then on the
return individuals of small-bodied fish species (\(TL < 10 cm\))\\
on a width of \(1 m\)
(see Triki et al.~2018 for further details on fish censuses data collection).
We then scaled the counts of cleaner fish, small-bodied,
and large-bodied clients fish densities per \(100 m^2\).
From the study by Triki \emph{et al} \citep{triki_Biological_2019}
cleaner fish and large-bodied client populations densities were
highly correlated, and only the former
was hence used in the analyses as representative of both measures.

@@

Previous research showed that cleaner fish living at high population
densities and giving service priority to the visitor plate in the
ephemeral reward task, as well as cleaner fish living at low
densities but denying service priority to the visitor plate possess
larger forebrains; a key teleost brain region associated with
behavioural flexibility and social intelligence. Those failing to
show optimized decision-rules given their local ecological
conditions had relatively smaller forebrains \citep{triki_Brain_2020}.
Triki \emph{et al.} refer to the former as socially competent cleaner fish,
while the second group as socially incompetent cleaner fish.
Social competence is the ability to optimise social behaviour
to the available social information \citep{taborsky_Social_2012, bshary_Cooperation_2015, varela_Correlated_2020}. Our analyses yielded no
evidence that the difference in social competence with respect to the local
ecological conditions and associated brain morphology,
found by Triki \emph{et al} \citep{triki_Brain_2020}, is due to the mechanism
used to incorporate future consequences. It is conceivable that
high performing individuals from low population densities reef sites
use negative reinforcement instead of chaining, but in that
case, negative reinforcement should have explained at least
part of the data. Configurational learning or
chunking \citep{sutherland_Configural_1989, miller_Magical_1956},
the second component necessary to solve the ephemeral
reward task \citep{quinones_Reinforcement_2019}, was
not varied in the models we analyzed here. However, while chunking tendencies
should vary to allow individuals to adapt to their local conditions
\citep{prat_Modelling_2022, kolodny_Evolution_2014}, systematic differences
in individual chunking tendencies would not explain how socially
competent decisions vary as a function of relative abundance.
Therefore, it remains currently unclear what cell-demanding
mechanisms may cause variation in social competence that
translates into site-specific variation in performance in the
ephemeral reward task.

>>>>

Over the last decade, our team tested over a hundred wild-caught
cleaner fish in the exact same paradigm of the ephemeral reward
task \citep{salwiczek_Adult_2012, wismer_Variation_2014, triki_Decrease_2018, triki_Biological_2019, triki_Brain_2020}. These fish often come from
different reef locations around the study field site at Lizard Island,
Great Barrier Reef in Australia. Further investigation of the local
eco-sociological conditions revealed that cleaner and client
fish population densities have a substantial impact on cleaner fish
performance in the task. Cleaner fish from reef sites with relatively
low densities were more likely to fail at solving the task \citep{triki_Biological_2019, triki_Decrease_2018, wismer_Variation_2014}. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply
low supply and high demand for cleaning services.
Under such conditions, there are fewer occasions under
which failing to give priority to a visitor translates to an empty
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities
are less likely to exert their partner choice options
and are hence more likely to wait for inspection if not given cleaning
priority \citep{triki_Brain_2020}. In the absence of partner choice
cleaner fish should not give priority to visitors any more.

@@

The empirical data were collected between 2010 and 2019 always during
the austral winter months June to August from a total of five
study reef sites (Corner Beach-CB, Horseshoe-HS, Mermaid Cove-MC,
Northern Horseshoe-NHS, and The Crest-TC) at Lizard Island
(\(14.6682^\circ S, 145.4604^\circ E\)), Great Barrier Reef, Australia.
The data consist of three sets: fish censuses, field
observations of cleaner-client interactions to quantify the probability
of visitors leaving if made to wait, and the
performance of wild-caught cleaner fish in the ephemeral reward test.
In total, we have twelve site/year data sets for fish censuses and
corresponding performance in the lab test. Thus, some sites were sampled more
than once. To estimate the population density of cleaner fish
and their clients at a given site in
a given year, Triki \emph{et al.} \citep{triki_Biological_2019} used a series of ten
transects of \(30m\) each. Observers swam along the transect lines placed on
the reef and first counted the visible large-bodied adult fish
(species with total length \(TL \geq 10cm\)) including cleaner fish
on a width of \(5m\), and then on the
return individuals of small-bodied fish species (\(TL < 10 cm\))\\
on a width of \(1 m\)
(see Triki et al.~2018 for further details on fish censuses data collection).
We then scaled the counts of cleaner fish, small-bodied,
and large-bodied clients fish densities per \(100 m^2\).
From the study by Triki \emph{et al} \citep{triki_Biological_2019}
cleaner fish and large-bodied client populations densities were
highly correlated, and only the former
was hence used in the analyses as representative of both measures.

@@

Previous research showed that cleaner fish living at high population
densities and giving service priority to the visitor plate in the
ephemeral reward task, as well as cleaner fish living at low
densities but denying service priority to the visitor plate possess
larger forebrains; a key teleost brain region associated with
behavioural flexibility and social intelligence. Those failing to
show optimized decision-rules given their local ecological
conditions had relatively smaller forebrains \citep{triki_Brain_2020}.
Triki \emph{et al.} refer to the former as socially competent cleaner fish,
while the second group as socially incompetent cleaner fish.
Social competence is the ability to optimise social behaviour
to the available social information \citep{taborsky_Social_2012, bshary_Cooperation_2015, varela_Correlated_2020}. Our analyses yielded no
evidence that the difference in social competence with respect to the local
ecological conditions and associated brain morphology,
found by Triki \emph{et al} \citep{triki_Brain_2020}, is due to the mechanism
used to incorporate future consequences. It is conceivable that
high performing individuals from low population densities reef sites
use negative reinforcement instead of chaining, but in that
case, negative reinforcement should have explained at least
part of the data. Configurational learning or
chunking \citep{sutherland_Configural_1989, miller_Magical_1956},
the second component necessary to solve the ephemeral
reward task \citep{quinones_Reinforcement_2019}, was
not varied in the models we analysed here. However, while chunking tendencies
should vary to allow individuals to adapt to their local conditions
\citep{prat_Modelling_2022, kolodny_Evolution_2014}, systematic differences
in individual chunking tendencies would not explain how socially
competent decisions vary as a function of relative abundance.
Therefore, it remains currently unclear what cell-demanding
mechanisms may cause variation in social competence that
translates into site-specific variation in performance in the
ephemeral reward task.
